---
title: "RWorksheet_Bernasol_4C"
author: "Shane Jhammielyn A. Bernasol"
output: pdf_document
---

#1: Importing and Exploring the `mpg` Dataset

```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(tidyr)

# Load the mpg dataset
data("mpg")

# Display the first few rows
head(mpg)
```

```{r}
#1B
categorical_vars <- mpg %>% select(where(is.character))
categorical_vars
```

#1C
```{r}
continuous_vars <- mpg %>% select(where(is.numeric))
continuous_vars

```


#2a. Group Manufacturers and Unique Models
```{r}
manufacturer_models <- mpg %>% 
  group_by(manufacturer) %>% 
  summarise(unique_models = list(unique(model)))

manufacturer_models
```

#2b. Graph Unique Models Using plot() and ggplot()
```{r}
# Count the models per manufacturer
manufacturer_counts <- mpg %>%
  group_by(manufacturer) %>%
  summarise(models = n_distinct(model))

# Plot the counts using ggplot
ggplot(manufacturer_counts, aes(x = reorder(manufacturer, -models), y = models)) +
  geom_bar(stat = "identity", fill = "pink") +
  theme_minimal() +
  labs(title = "Number of Models per Manufacturer", 
       x = "Manufacturer", 
       y = "Model Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# Using ggplot2
ggplot(manufacturer_counts, aes(x = reorder(manufacturer, -models), y = models)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(title = "Unique Models per Manufacturer (ggplot2)",
       x = "Manufacturer",
       y = "Number of Unique Models") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

#2. What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?
```{r}
ggplot(mpg, aes(x = model, y = manufacturer)) + geom_point()

```


#2a. This plot shows the relationship between model and manufacturer. However, it may not be very informative due to the large number of model names, which can overlap and create a cluttered appearance, making the plot difficult to interpret.

#3. Plot the model and the year using ggplot(). Use only the top 20 observations.
```{r}
# Get the top 20 observations based on year
top_20_mpg <- mpg %>% 
  arrange(desc(year)) %>% 
  head(20)

# Create the plot
ggplot(top_20_mpg, aes(x = model, y = year)) +
  geom_point(color = "purple") +
  labs(title = "Top 20 Models by Year",
       x = "Model",
       y = "Year") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
#4a. Plot using geom_bar() using the top 20 observations only. The graphs should have a title, labels, and colors.
```{r}
# Group the data by model and count the number of cars per model
model_counts <- mpg %>%
  group_by(model) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  head(20)

# Plot the top 20 models with geom_bar()
ggplot(model_counts, aes(x = reorder(model, -count), y = count, fill = count)) +
  geom_bar(stat = "identity") +
  labs(title = "Top 20 Car Models by Count",
       x = "Car Model",
       y = "Number of Cars",
       fill = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_gradient(low = "pink", high = "red")

```
#4b. Plot using the geom_bar() + coord_flip().
```{r}
# Plot the top 20 models with geom_bar() and coord_flip()
ggplot(model_counts, aes(x = reorder(model, -count), y = count, fill = count)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top 20 Car Models by Count",
       x = "Car Model",
       y = "Number of Cars",
       fill = "Count") +
  theme_minimal() +
  scale_fill_gradient(low = "pink", high = "red")

```
#5. Plot the relationship between cyl - number of cylinders and displ - engine displacement using geom_point with aesthetic color = engine displacement.
```{r}
# Create the plot
ggplot(mpg, aes(x = cyl, y = displ, color = displ)) +
  geom_point() +
  labs(title = "Relationship between No. of Cylinders and Engine Displacement",
       x = "Number of Cylinders",
       y = "Engine Displacement (L)") +
  theme_minimal() +
  scale_color_gradient(low = "pink", high = "red")

```

6.
```{r}
library(ggplot2)

mpg_data <- ggplot2::mpg

ggplot(mpg_data, aes(x = displ, y = hwy, color = cty)) +
  geom_point() +
  labs(
    title = "Relationship between Engine Displacement and Highway MPG",
    x = "Engine Displacement (L)",
    y = "Highway Miles per Gallon",
    color = "City MPG"
  ) +
  scale_color_gradient(low = "pink", high = "red")
```
6a.
```{r}
#6a. Display observations and structure of the traffic dataset
traffic_data <- read.csv("/cloud/project/Worksheet 4/Worksheet 4C/traffic.csv")

# View column names
colnames(traffic_data)

# View the structure of the dataset
str(traffic_data)

# Display the first few rows to check for formatting issues
head(traffic_data)


```
6b.
```{r}
junctions <- split(traffic_data, traffic_data$Junction)
junctions
```

6c.
```{r}
traffic_data$DateTime <- as.POSIXct(traffic_data$DateTime, format = "%Y-%m-%d %H:%M:%S")
ggplot(traffic_data, aes(x = DateTime, y = Vehicles)) +
geom_line(color = "pink") +
labs(title = "Traffic Volume Over Time at Each Junction",
x = "DateTime",
y = "Number of Vehicles") +
facet_wrap(~ Junction)

```
7a.
```{r}
library(readxl)

alexa_data <- read_excel("/cloud/project/Worksheet 4/Worksheet 4C/alexa_file.xlsx")

str(alexa_data)

```
```{r}
ncol(alexa_data)
```

7b.
```{r}
library(dplyr)
print(colnames(alexa_data))
```
7b.
```{r}
variation_totals <- alexa_data %>%
  group_by(variation) %>%
  summarize(total = n())

variation_totals

```
7c.
```{r}
library("ggplot2")
ggplot(alexa_data, aes(x = date, y = variation)) +
  geom_point() +
  labs(title = "Variations Timeline")

```

#6a. How many numbers of observation does it have? What are the variables of the traffic dataset?



