---
title: "RWorksheet_Bernasol#3A"
author: "Shane Jhammielyn A. Bernasol"
output: pdf_document
---

# a. First 11 letters
```{r}
first11 <- LETTERS[1:11]
print(first11)
```


# b. Odd-numbered letters
```{r}
oddLetters <- LETTERS[seq(1, 26, by = 2)]
print(oddLetters)
```


# c. vowels
```{r}
vowels <- LETTERS[c(1, 5, 9, 15, 21)]
print(vowels)
```


# d. Last 5 lowercase letters
```{r}
last5 <- letters[22:26]
print(last5)
```


# e. Lowercase letters between 15 to 24
```{r}
midLetters <- letters[15:24]
print(midLetters)
```


# a. Create character vector for city names
```{r}
city <- c("Tuguegarao City", "Manila", "Iloilo City", "Tacloban", "Samal Island", "Davao City")
print(city)
```


# b. Create temperature vector
```{r}
temp <- c(42, 39, 34, 34, 30, 27)
print(temp)
```

# c. Create a dataframe combining city and temp
```{r}
city_temp <- data.frame(City = city, Temperature = temp)
print(city_temp)
```


# d. Renaming the columns of the dataframe
```{r}
names(city_temp) <- c("City", "Temperature")
print(city_temp)
```


# e. Printing the structure of the dataframe
```{r}
str(city_temp)
```


# f. Content of row 3 and row 4
```{r}
print(city_temp[3:4, ])
```


# g. City with the highest and lowest temperature
```{r}
hottest <- city_temp[which.max(city_temp$Temperature), ]
coldest <- city_temp[which.min(city_temp$Temperature), ]
print(hottest)
print(coldest)
```


# a. Create a matrix with 4 columns and 3 rows
```{r}
mat <- matrix(c(1:8, 11:14), nrow = 3, ncol = 4)
print(mat)
```


# b. Multiply the matrix by 2
```{r}
mat2 <- mat * 2
print(mat2)
```


# c. Content of row 2
```{r}
print(mat2[2, ])
```


# d. Column 3 and 4 from row 1 and 2
```{r}
print(mat2[1:2, 3:4])
```


# e. Columns 2 and 3 from row 3
```{r}
print(mat2[3, 2:3])
```


# f. Only column 4
```{r}
print(mat2[, 4])
```

# g. Renaming rows and columns
```{r}
rownames(mat2) <- c("isa", "dalawa", "tatlo")
colnames(mat2) <- c("uno", "dos", "tres", "quatro")
print(mat2)
```


# h. Reshape the matrix into 2 columns and 6 rows
```{r}
dim(mat) <- c(6, 2)
print(mat)
```


# a. Create a 3D array with specified values, 4 columns, 2 rows, and 3 dimensions
```{r}
arr <- array(rep(c(1, 2, 3, 6, 7, 8, 9, 0, 3, 4, 5, 1), each = 2), dim = c(2, 4, 3))
print(arr)
```

# b. Checking the dimensions of the array
```{r}
print(dim(arr))
```


# c. Naming the dimensions of the array
```{r}
dimnames(arr) <- list(letters[1:2], LETTERS[1:4], c("1st-Dimensional Array", "2nd-Dimensional Array", "3rd-Dimensional Array"))
print(arr)
```


